---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

1. Load the data

```{r,echo=TRUE}
activity<-read.csv("activity.csv")
```

2. Process/transform the data (if necessary) into a format suitable for your analysis

Note: date is provided as factor.  Need to transform it to date format.

```{r,echo=TRUE}
activity$date<-as.Date(activity$date)
```


##What is mean total number of steps taken per day?

Ignoring the missing values in the dataset:

1. Calculate the total number of steps taken per day

```{r, echo=TRUE}
total_steps<-tapply(activity$steps,activity$date,sum, rm.na=TRUE)
```


2. Make a histogram of the total number of steps taken each day

```{r, echo=TRUE}
hist(total_steps, main = "Histogram of steps taken each day", xlab = "steps per day", axes = TRUE, breaks=10)
```


3. Calculate and report the mean and median of the total number of steps taken per day

```{r, echo=TRUE}
mean_steps<-summary(tapply(activity$steps,activity$date,sum, rm.na=TRUE))[4]
median_steps<-summary(tapply(activity$steps,activity$date,sum, rm.na=TRUE))[3]
```

The mean number of steps taken by day is `r sprintf("%i" , mean_steps)` 

The median number of steps taken by day is `r sprintf("%i" , median_steps)` 


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
